---
title: "SGG - Who are the main player?"
author: "André Soares"
date: "8 January 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r initial loadings, include=FALSE}
library(dada2); packageVersion("dada2")
library(ggplot2); packageVersion("ggplot2")
library(phyloseq); packageVersion("phyloseq")
library(data.table); packageVersion("data.table")
library(DESeq2); packageVersion("DESeq2")
library(dplyr)
library(plyr)
library(magrittr)
library(RColorBrewer)
library(phangorn)
library(DECIPHER)
library("structSSI")
library(ips)
library(msa)
library(picante)
library(igraph)
library(cowplot)
library(plotly)
library(lemon)
```

```{r load SV data, echo=FALSE}
SGG_SVt = readRDS("/media/andre/B2F8C9A0F8C962E9/SGG_16S_analysis/SGG_16S_data_trimmed_DADA2_table/seqtab_final.rds")
SGG_Tax = readRDS("/media/andre/B2F8C9A0F8C962E9/SGG_16S_analysis/SGG_16S_data_trimmed_DADA2_table/tax_final.rds")
sgg_metadata<-read.csv("/media/andre/B2F8C9A0F8C962E9/SGG_16S_analysis/SGG_16S_metadata/SGG_metadata.csv", header=TRUE)
order<-rownames(SGG_SVt)
sgg_metadata_ord<-sgg_metadata[match(order, sgg_metadata$Sample_Code),]
rownames(sgg_metadata_ord) <- sgg_metadata_ord$Sample_Code
ps <- phyloseq(otu_table(SGG_SVt, taxa_are_rows=FALSE),
               sample_data(sgg_metadata_ord), 
               tax_table(SGG_Tax))
ps.a = subset_samples(ps, Site.name != "Taffs Well PUMPED")
ps.b = subset_samples(ps.a, Sample_type != "Control")
ps.b <- prune_taxa(taxa_sums(ps.b) > 0, ps.b)
#order months
sample_data(ps.b)$Month = factor(sample_data(ps.b)$Month, 
                                 levels = c("April","August","December"))
#to relative abundances
ps.b = subset_taxa(ps.b, Kingdom %in% c("Archaea", "Bacteria"))
ps.b.r <-  transform_sample_counts(ps.b, function(x) {x/sum(x)} ) 
```

##1 - General view of taxonomic profiles across the dataset

```{r general proteobacterial taxonomy}
# agglomerate taxa
ps.b.r.glom <- tax_glom(ps.b.r, taxrank = 'Phylum', NArm = FALSE)
# create dataframe from phyloseq object
ps.b.r.glom.psdf <- data.table(psmelt(ps.b.r.glom))
# convert Phylum to a character vector from a factor because R
ps.b.r.glom.psdf$Phylum <- as.character(ps.b.r.glom.psdf$Phylum)
# group dataframe by Phylum, calculate median rel. abundance
ps.b.r.glom.psdf[, median := median(Abundance, na.rm = TRUE), 
                 by = "Phylum"]
# Change name to remainder of Phylum less than 1%
ps.b.r.glom.psdf[(median <= 0.01), Phylum := "Other"]
#Removing Proteobacteria from previous table
ps.b.r.glom.psdf.noProteo<-ps.b.r.glom.psdf[!grepl("Proteobacteria", ps.b.r.glom.psdf$Phylum),]
#New table, with Proteobacterial classes only
ps.b.r.ProtCl = subset_taxa(ps.b.r, Phylum == "Proteobacteria")
ps.b.r.ProtCl.glom <- tax_glom(ps.b.r.ProtCl, taxrank = 'Class', NArm = FALSE)
# create dataframe from phyloseq object
ps.b.r.ProtCl.glom.psdf <- data.table(psmelt(ps.b.r.ProtCl.glom))
# convert Class to a character vector from a factor because R
ps.b.r.ProtCl.glom.psdf$Class <- as.character(ps.b.r.ProtCl.glom.psdf$Class)
# group dataframe by Phylum, calculate median rel. abundance
ps.b.r.ProtCl.glom.psdf[, median := median(Abundance, na.rm = TRUE), 
                        by = "Class"]
# Change name to remainder of Phylum less than 1%
ps.b.r.ProtCl.glom.psdf[(median <= 0.01), Class := "Other Proteobacteria"]
ps.b.r.ProtCl.glom.psdf.noP <- subset(ps.b.r.ProtCl.glom.psdf, select = -Phylum)
#same name= allows merging
names(ps.b.r.glom.psdf.noProteo)[names(ps.b.r.glom.psdf.noProteo) == 'Phylum'] <- 'Taxa'
names(ps.b.r.ProtCl.glom.psdf.noP)[names(ps.b.r.ProtCl.glom.psdf.noP) == 'Class'] <- 'Taxa'
#join the two
ps.b.r.ProteoCl.AllPhy <- rbind(ps.b.r.glom.psdf.noProteo, ps.b.r.ProtCl.glom.psdf.noP)
ps.b.r.ProteoCl.AllPhy.plot<-ggplot(ps.b.r.ProteoCl.AllPhy[Abundance > 0], 
                                    aes(x = Site.name, y = Abundance/3, fill = Taxa)) + 
  facet_grid(Month~.) +
  geom_bar(stat = "identity") +
  # Remove x axis title
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  guides(fill = guide_legend(keywidth = 1, keyheight = 1)) +
  # coord_cartesian(ylim = c(0, 1)) +
  ylab("Relative Abundance\n")
#totals vary because the bars = medians
```
```{r , fig.height=14, fig.width=20}
ps.b.r.ProteoCl.AllPhy.plot
```

Clearly, the dataset is mostly populated by Beta- and Epsilonproteobacteria, which seem to correspond to iron- and sulphur-rich sites, respectively.

Is this dominance explain by particular genera? Yep, seems like so. Also, some temporal trends seem to pop up. We'll take a look at that later.

```{r genera plot}
ps.b.r.HelicobGall = subset_taxa(ps.b.r, Genus=="Sulfuricurvum" | Genus=="Sulfurimonas" | Genus=="Sulfurovum" | Genus=="Gallionella" | Genus=="Sideroxydans" | Genus=="Ferriphaselus")
ps.b.r.HelicobGall.glom <- tax_glom(ps.b.r.HelicobGall, taxrank = 'Genus', NArm = FALSE)
ps.b.r.HelicobGall.glom.psdf <- data.table(psmelt(ps.b.r.HelicobGall.glom))
ps.b.r.HelicobGall.glom.psdf <- data.frame(lapply(ps.b.r.HelicobGall.glom.psdf, function(x) {gsub("Crumlin Navigation", "Crum. Nav.", x)}))
ps.b.r.HelicobGall.glom.psdf$Genus <- as.character(ps.b.r.HelicobGall.glom.psdf$Genus)
ps.b.r.HelicobGall.glom.psdf$Abundance <- as.numeric(as.character(ps.b.r.HelicobGall.glom.psdf$Abundance))

ps.b.r.HelicobGall.glom.psdf.plot<-ggplot(ps.b.r.HelicobGall.glom.psdf,
                                          aes(x=Genus,y=Abundance,fill=Genus)) + 
  geom_point(aes(color=Genus, fill=Genus,
                 size=Abundance)) +
  geom_smooth(method="loess") +
  facet_grid(Site.name~Month)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text.y = element_text(angle = 0)) +
  coord_fixed(ratio = 2)
```
```{r , fig.height=20, fig.width=14}
ps.b.r.HelicobGall.glom.psdf.plot
```

It seems though from this and the latter graph that Lindsay has a much more diverse profile i.e. more than 40% of the profile is made up of 'Other' phyla. Let's take a closer look at that.

```{r lindsay}
ps.b.r.LindsayCHen = subset_samples(ps.b.r, Site.name=="Lindsay")
ps.b.r.glom.phy <- tax_glom(ps.b.r.LindsayCHen, taxrank = 'Phylum', NArm = FALSE )
ps.b.r.glom.phy.psdf <- data.table(psmelt(ps.b.r.glom.phy))
ps.b.r.glom.phy.psdf.f <- ps.b.r.glom.phy.psdf[ , c("Abundance","Phylum")]
ps.b.r.glom.phy.psdf.f.s<-ps.b.r.glom.phy.psdf.f[, .(count=.N, var=mean(Abundance)), by = Phylum]

ps.b.r.glom.phy.psdf.f.s <- transform(ps.b.r.glom.phy.psdf.f.s, Phylum=reorder(Phylum, -var)) 
ps.b.r.glom.phy.psdf.f.s.plot=ggplot(ps.b.r.glom.phy.psdf.f.s, aes(x=Phylum, y=var, color=Phylum)) +
  geom_point() +
  geom_hline(yintercept = 0.01,
             color="gray",
             linetype="dashed") +
  ylab("Relative abundance") +
  theme(axis.text.x = element_text(angle=40, hjust = 1, vjust = 1),
        axis.title.x = element_blank(),
        legend.position = "none")
```
```{r ,fig.width=14}
ps.b.r.glom.phy.psdf.f.s.plot
```

This site is clearly populated by a microbial community where a number of phyla co-exist in relatively low (between 25 and 5%) proportions.

So, how does this translate to beta-diversity?

##2 - Community structure

```{r pcoa}
ps.b.r.pcoa  <- ordinate(ps.b.r, "PCoA", distance="jsd", 
                         k=2, trymax=1e3, weighted=TRUE)
evals<-ps.b.r.pcoa$values$Eigenvalues
#Axis 1 and 2
pcoa.a12<-sum(evals[1:2])/sum(evals)
pcoa.a12.r<-round(pcoa.a12, digits=4)*100
#Axis 1 and 3
pcoa.a13<-sum(evals[c(1,3)])/sum(evals)
pcoa.a13.r<-round(pcoa.a13, digits=4)*100
ps.b.r.pcoa.plot.a1 <- plot_ordination(ps.b.r, ps.b.r.pcoa, axes = c(1,2),
                                       color="Site.name", shape="Month") +
  geom_polygon(aes(fill=Site.name), alpha=0.5) +
  geom_point(alpha=0.5, size=4)+
  scale_alpha(guide = 'none') +
  theme(legend.position = "none") +
  annotate("text", y= 0.3, x =0.25,label=paste("Axes 1 and 2:\n",pcoa.a12.r,"% variation")
           ,hjust=0.5, size=5)

ps.b.r.pcoa.plot.a2 <- plot_ordination(ps.b.r, ps.b.r.pcoa, axes = c(1,3),
                                       color="Site.name", shape="Month")+
  geom_polygon(aes(fill=Site.name), alpha=0.5) +
  geom_point(alpha=0.5, size=4) +
  scale_alpha(guide = 'none')+
  annotate("text", y= 0.2, x =0.25,label=paste("Axes 1 and 3:\n",pcoa.a13.r,"% variation")
           ,hjust=0.5, size=5)

ps.b.r.pcoa.plot<-plot_grid(ps.b.r.pcoa.plot.a1, ps.b.r.pcoa.plot.a2, 
                            labels=c("A","B"))
ps.b.r.pcoa.plot
```
```{r , fig.height=18, fig.width=20}
ps.b.r.pcoa.plot
```

Ok then, a clear separation between iron- and sulphur-rich sites.

Not enough resolution to infer year-long patterns, so let's take a closer look at how beta-diversity varies across time.

##3 - SGG across time

```{r pcoa axis1 month}
ps.b.r.pcoa.v = data.table(ps.b.r.pcoa$vectors, keep.rownames = TRUE)
sdt = data.table(as(sample_data(ps.b.r), "data.frame"), keep.rownames = TRUE)

setnames(ps.b.r.pcoa.v, "rn", "Sample.ID")
df <- subset(sdt, select = -c(Sample.ID) )
setnames(df, "rn", "Sample.ID")
setkey(ps.b.r.pcoa.v, Sample.ID)
setkey(df, Sample.ID)
ps.b.r.pcoa.v <- ps.b.r.pcoa.v[df]
# Axis 1
ps.b.r.pcoa.v.plot=ggplot(ps.b.r.pcoa.v, aes(Month, Axis.1, color = Site.name)) +
  geom_point(size = 3) +
  geom_point(size = 5, alpha=0.4) + 
  facet_wrap(~Site.name) +
  theme(axis.text.x = element_text(angle = 25, vjust=1, hjust=1)) +
  ylab("JSD PCoA Axis 1")
```
```{r , fig.height=12, fig.width=14}
ps.b.r.pcoa.v.plot
```

It seems that except for month-specific outliers, seasonality doesn't have much effect on the beta-diversity.

Is alpha-diversity impacted by seasonality?

```{r alpha div}
ps.b.rich=plot_richness(ps.b, x="Month", measures="Shannon",
                        color="Site.name") + 
  geom_point(size = 5, alpha=0.4) +
  facet_wrap(~Site.name) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 25, vjust=1, hjust=1),
        axis.title.x = element_blank()) +
  ylab("Shannon Diversity")
```
```{r , fig.height=12, fig.width=14}
ps.b.rich
```

It seems like so. Specifically, Blaenavon and Celynen N seem to decrease along the year. On the other side, Cefn Hengoed, Glyncastle, Morlais seem to be positively correlated with the passing of the year.

```{r investigate neg pos cors month}
ps.b.r.posnegs = subset_samples(ps.b.r, Site.name=="Blaenavon" | Site.name=="Celynen North" | Site.name=="Cefn Hengoed" | Site.name=="Glyncastle" | Site.name=="Morlais")
ps.b.r.posnegs.class = subset_taxa(ps.b.r.posnegs, Class=="Betaproteobacteria" | Class=="Deltaproteobacteria" | Class=="Gammaproteobacteria" | Class=="Epsilonproteobacteria" | Class=="Elusimicrobia")
ps.b.r.posnegs.glom <- tax_glom(ps.b.r.posnegs.class, taxrank = 'Order', NArm = FALSE)
ps.b.r.posnegs.glom.psdf <- data.table(psmelt(ps.b.r.posnegs.glom))
ps.b.r.posnegs.glom.psdf$Order <- as.character(ps.b.r.posnegs.glom.psdf$Order)
ps.b.r.posnegs.glom.psdf$Abundance <- as.numeric(ps.b.r.posnegs.glom.psdf$Abundance)
ps.b.r.posnegs.glom.psdf.plot<-ggplot(ps.b.r.posnegs.glom.psdf[Abundance > 0.05],
                                          aes(x=Order,y=Abundance,
                                              fill=Order)) + 
  geom_point(aes(color=Order, fill=Order,
                 size=Abundance)) + 
  facet_rep_grid(Site.name~Month)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  coord_fixed(ratio = 6)
```
```{r , fig.height=12, fig.width=14}
ps.b.r.posnegs.glom.psdf.plot
```

Ok then, some clear trends showing up, let's try and identify if this is a large-scale effort or the work of specific SVs.

```{r investigate neg pos cors month deeper}
ps.b.r.posnegs.genus.glom <- tax_glom(ps.b.r.posnegs.class, taxrank = 'Genus', NArm = FALSE)
ps.b.r.posnegs.genus.glom.psdf <- data.table(psmelt(ps.b.r.posnegs.genus.glom))
ps.b.r.posnegs.genus.glom.psdf$Genus <- as.character(ps.b.r.posnegs.genus.glom.psdf$Genus)
ps.b.r.posnegs.genus.glom.psdf$Abundance <- as.numeric(ps.b.r.posnegs.genus.glom.psdf$Abundance)
ps.b.r.posnegs.genus.glom.psdf.plot<-ggplot(ps.b.r.posnegs.genus.glom.psdf[Abundance > 0.01],
                                          aes(x=Genus,y=Abundance,
                                              fill=Genus)) + 
  geom_point(aes(color=Genus, fill=Genus,
                 size=Abundance)) +
  facet_rep_grid(Site.name~Month) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  guides(col = guide_legend(nrow = 22))
```
```{r , fig.height=12, fig.width=14}
ps.b.r.posnegs.genus.glom.psdf.plot
```

Interesting stuff, seems Rhodoferax was ignored as a significant contributor for Betaproteobacterial trends across the dataset. Revisit that.
