---
title: "SGG - How does metadata relate to the microbiome?"
author: "André Soares"
date: "8 January 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r initial loadings, include=FALSE}
library(dada2); packageVersion("dada2")
library(ggplot2); packageVersion("ggplot2")
library(phyloseq); packageVersion("phyloseq")
library(data.table); packageVersion("data.table")
library(DESeq2); packageVersion("DESeq2")
library(dplyr)
library(plyr)
library(magrittr)
library(RColorBrewer)
library(phangorn)
library(DECIPHER)
library("structSSI")
library(ips)
library(msa)
library(picante)
library(igraph)
library(cowplot)
library(plotly)
library(lemon)
library(GGally)
```

```{r load SV data, echo=FALSE}
SGG_SVt = readRDS("/media/andre/B2F8C9A0F8C962E9/SGG_16S_analysis/SGG_16S_data_trimmed_DADA2_table/seqtab_final.rds")
SGG_Tax = readRDS("/media/andre/B2F8C9A0F8C962E9/SGG_16S_analysis/SGG_16S_data_trimmed_DADA2_table/tax_final.rds")
sgg_metadata<-read.csv("/media/andre/B2F8C9A0F8C962E9/SGG_16S_analysis/SGG_16S_metadata/SGG_metadata.csv", header=TRUE)
order<-rownames(SGG_SVt)
sgg_metadata_ord<-sgg_metadata[match(order, sgg_metadata$Sample_Code),]
rownames(sgg_metadata_ord) <- sgg_metadata_ord$Sample_Code
ps <- phyloseq(otu_table(SGG_SVt, taxa_are_rows=FALSE),
               sample_data(sgg_metadata_ord), 
               tax_table(SGG_Tax))
ps.a = subset_samples(ps, Site.name != "Taffs Well PUMPED")
ps.b = subset_samples(ps.a, Sample_type != "Control")
ps.b <- prune_taxa(taxa_sums(ps.b) > 0, ps.b)
#order months
sample_data(ps.b)$Month = factor(sample_data(ps.b)$Month, 
                                 levels = c("April","August","December"))
#to relative abundances
ps.b = subset_taxa(ps.b, Kingdom %in% c("Archaea", "Bacteria"))
ps.b.r <-  transform_sample_counts(ps.b, function(x) {x/sum(x)} ) 
```

##1 - Visualizing SGG metadata


```{r main metadata}
sample_data(ps.b.r)$Site.name = factor(sample_data(ps.b.r)$Site.name, 
                                          levels = c("Blaenavon","Cefn Hengoed","Mountain Gate","Taff Bargoed",
                                                     "Glyncastle","Dinas","Ynysarwed","Celynen North","Morlais",
                                                     "Lindsay","Six Bells","Crumlin Navigation","Taffs Well"))
metadata<-sample_data(ps.b.r)
#Temp
temp_table.plot<-ggplot(metadata, 
                        aes(x=Site.name, y=Temp.C, fill=Month)) +
  geom_point(aes(fill=Month, color=Month),
             size=2,
             position = position_dodge(width=0.5)) +
  geom_hline(yintercept = 13.4,
             color="dark red",
             linetype="dashed") +
  geom_hline(yintercept = 11,
             color="gray",
             linetype="dashed") +
  theme(axis.title.x=element_blank(),
        axis.title.y = element_text(size=14),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size=12),
        legend.position = "none") +
  annotate("text", "Mountain Gate", 14, vjust = -1, hjust=.45, size= 4,
           label = "Average mine water temperature (Farr et al, 2016)") +
  labs(y=expression(paste("Temperature (",degree,"C)")))

#DO
do.plot<-ggplot(metadata, 
                        aes(x=Site.name, y=`DO..sat`, fill=Month)) +
  geom_point(aes(fill=Month, color=Month),
             size=2,
             position = position_dodge(width=0.5)) +
  geom_line(aes(group=Month, color=Month),
            stat="smooth", method="loess", 
            alpha = 0.4, size=1.5) +
  geom_hline(yintercept = 50,
           color="gray60",
           linetype="dashed") +
  theme(axis.title.x=element_blank(),
        axis.title.y = element_text(size=14),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size=12),
        legend.position = "none") +
  labs(y="Dissolved Oxygen (%)")
#Sulfate
so4.plot<-ggplot(metadata, 
                        aes(x=Site.name, y=`SO4..µg.ml`, fill=Month)) +
  geom_point(aes(fill=Month, color=Month),
             size=2,
             position = position_dodge(width=0.5)) +
  geom_line(aes(group=Month, color=Month),
            stat="smooth", method="loess", 
            alpha = 0.4, size=1.5) +
  geom_hline(yintercept = 500,
         color="gray60",
         linetype="dashed") +
  theme(axis.title.x=element_blank(),
        axis.title.y = element_text(size=14),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size=12),
        legend.position = "none") +
  labs(y=expression(paste("SO"[4]^{-2},"(",mu,"g.mL"^{-1},")"))) +
  guides(size=FALSE)
#pH
pH.plot<-ggplot(metadata, 
                        aes(x=Site.name, y=`pH`, fill=Month)) +
  geom_point(aes(fill=Month, color=Month),
             size=2,
             position = position_dodge(width=0.5)) +
  geom_line(aes(group=Month, color=Month),
            stat="smooth", method="loess", 
            alpha = 0.4, size=1.5) +
  geom_hline(yintercept = 7,
           color="gray60",
           linetype="dashed") +
  theme(axis.title.x=element_blank(),
        axis.title.y = element_text(size=14),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size=12),
        legend.position = "none") +
  labs(y="pH") +
  guides(size=FALSE)
#EC
ec.plot<-ggplot(metadata, 
                        aes(x=Site.name, y=`EC.uS.cm`, fill=Month)) +
  geom_point(aes(fill=Month, color=Month),
             size=2,
             position = position_dodge(width=0.5)) +
  geom_line(aes(group=Month, color=Month),
            stat="smooth", method="loess", 
            alpha = 0.4, size=1.5) +
  geom_hline(yintercept = 2000,
           color="gray60",
           linetype="dashed") +
  geom_hline(yintercept = 1000,
           color="grey80",
           linetype="dashed") +
  scale_y_continuous(breaks=c(0,500,1000,1500,2000,2500,3000)) +
  theme(axis.title.x=element_blank(),
        axis.title.y = element_text(size=14),
        axis.text.x = element_text(angle = 25, hjust = 1, size= 14),
        axis.text.y = element_text(size=12),
        legend.position = "top",
        legend.title = element_text(face = "bold"),
        legend.background = element_rect(fill="gray95", size=.5, linetype="dotted")) +
  labs(y=expression(paste("EC (",mu,"S.cm"^{-1},")")))
```
```{r message=FALSE, warning=FALSE, fig.height=11, fig.width=10}
plot_grid(temp_table.plot, do.plot, so4.plot, pH.plot, ec.plot, ncol = 1, align='v', rel_heights = c(1,1,1,1,1.75))
```

Coolsies, but these are only 5 of them. How does the overal picture look?

```{r metadata cors}
metadata_df=as(sample_data(ps.b.r),"data.frame")
#remove all columns with one or more NA entry
#NA in DO substituted to the mean of the two values in the site
metadata_df$DO..sat[is.na(metadata_df$DO..sat)] <- 44.1
metadata_noNA = metadata_df[ , colSums(is.na(metadata_df)) == 0]
cols_to_go <- names(metadata_noNA) %in% c("Sample_Code", "Sample.ID", "Sample_type","Site.ID","Site.name","Time_point","Month","Batch",
                               "Date","Time")
metadata_noNA_f <- metadata_noNA[!cols_to_go]
```
```{r message=FALSE, warning=FALSE, fig.height=11, fig.width=12}
ggcorr(metadata_noNA_f, method = c("everything", "kendall"), 
       nbreaks = 10, hjust = 1, size=4, layout.exp = 2)
```

There's clearly a lot of variation, but definitely there's loads more to learn on how these variables relate.

But how do they relate to microbial community structure?

```{r ordination metadata}
ps_sig <- phyloseq(otu_table(SGG_SVt, taxa_are_rows=FALSE),
                   sample_data(metadata_noNA), 
                   tax_table(SGG_Tax))
sample_data(ps_sig)$Month = factor(sample_data(ps_sig)$Month, 
                                   levels = c("April","August","December","Control"))
ps_sig_ntf = subset_samples(ps_sig, Site.name != "Taffs Well PUMPED")
ps_sig_ntf_nc = subset_samples(ps_sig_ntf, Sample_type != "Control")
ps_sig_ntf_nc_noab = subset_taxa(ps_sig_ntf_nc, Kingdom %in% c("Archaea", "Bacteria"))
ps_cap_r = transform_sample_counts(ps_sig_ntf_nc_noab, function(x) x/sum(x))

set.seed(1)
# Calculate Jensen-Shannon distance matrix
ps_cap_r_jsd <- phyloseq::distance(ps_cap_r, method = "jsd")
# make a data frame from the sample_data
sampledf_cap <- data.frame(sample_data(ps_cap_r))
# Adonis test: "how is variation attributed to different experimental" variables?
# null hypothesis: tested metadata categories have the same centroid.
adonis(ps_cap_r_jsd ~ Site.name + Temp.C + Month + EC.uS.cm + DO..sat + K.µg.ml + Na.µg.ml + Cl..µg.ml + SO4..µg.ml + X6.Li.ng.ml + X7.Li.ng.ml + X7.Li..He..ng.ml + X11.B.ng.ml + X45.Sc.ng.ml + X55.Mn.ng.ml + X59.Co.ng.ml + X60.Ni.ng.ml + X69.Ga.ng.ml + X75.As..He..ng.ml + X78.Se.ng.ml + X85.Rb.ng.ml + X88.Sr.ng.ml + X89.Y.ng.ml + X125.Te.ng.ml + X133.Cs.ng.ml + X135.Ba.ng.ml + X139.La.ng.ml + X146.Nd.ng.ml + X153.Eu.ng.ml + X166.Er.ng.ml + X175.Lu.ng.ml + X238.U.ng.ml, 
       data = sampledf_cap, strata = sampledf_cap$Time_point, permutations=1e3)
```

Yay, so much crap! Let's keep the significant variables and include them in the CAP formula.

```{r cap on sig metadata, message=FALSE, warning=FALSE}
#CAP
CAP_ps_cap_r <- ordinate(ps_cap_r,
                           method = "CAP",
                           distance = "jsd",
                           k=3, trymax=1e3, weighted=TRUE,
                           formula = ~ Site.name + Month + Temp.C + DO..sat + Cl..µg.ml + SO4..µg.ml+ X55.Mn.ng.ml + X89.Y.ng.ml + X139.La.ng.ml)

CAP_sig_plot_a12 <- plot_ordination(ps_cap_r,
                                    ordination = CAP_ps_cap_r, color = "Site.name", axes = c(1,2)) +
  aes(shape = Month) +
  geom_point(aes(colour = Site.name), alpha = 0.4, size = 4) +
  geom_polygon(aes(fill=Site.name), alpha=0.5) +
  geom_point(aes(colour = Site.name), size = 1.5)+
  scale_color_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(13)) +
  scale_fill_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(13)) +
  theme(legend.position = "none")

arrowmat_CAP <- vegan::scores(CAP_ps_cap_r, display = "bp")
arrowmat_CAP<-arrowmat_CAP[!grepl("Site.name", rownames(arrowmat_CAP)),]
arrowmat_CAP<-arrowmat_CAP[!grepl("Month", rownames(arrowmat_CAP)),]
arrowdf_CAP <- data.frame(labels = rownames(arrowmat_CAP), arrowmat_CAP)
arrow_map_CAP <- aes(xend = CAP1, yend = CAP2,
                     x = 0, y = 0,
                     shape = NULL, color = NULL,
                     label = labels)
label_map_CAP <- aes(x = 1.3 * CAP1, y = 1.3 * CAP2,
                     shape = NULL, color = NULL,
                     label = labels)
arrowhead_CAP = arrow(length = unit(0.02, "npc"))

CAP_sig_plot_a12 = CAP_sig_plot_a12 +
  geom_segment(mapping = arrow_map_CAP, size = .5, 
               data = arrowdf_CAP, color = "gray", arrow = arrowhead_CAP) +
  geom_text(mapping = label_map_CAP, size = 4, 
            data = arrowdf_CAP, show.legend = FALSE)

arrowmat_CAP_13 <- vegan::scores(CAP_ps_cap_r, display = "bp", choices=c(1,3))
arrowmat_CAP_13<-arrowmat_CAP_13[!grepl("Site.name", rownames(arrowmat_CAP_13)),]
arrowmat_CAP_13<-arrowmat_CAP_13[!grepl("Month", rownames(arrowmat_CAP_13)),]
arrowdf_CAP_13 <- data.frame(labels = rownames(arrowmat_CAP_13), arrowmat_CAP_13)
arrow_map_CAP_13 <- aes(xend = CAP1, yend = CAP3,
                        x = 0, y = 0,
                        shape = NULL, color = NULL,
                        label = labels)
label_map_CAP_13 <- aes(x = 1.3 * CAP1, y = 1.3 * CAP3,
                        shape = NULL, color = NULL,
                        label = labels)
arrowhead_CAP_13 = arrow(length = unit(0.02, "npc"))

CAP_sig_plot_a13 <- plot_ordination(ps_cap_r,
                                    ordination = CAP_ps_cap_r, color = "Site.name",axes = c(1,3)) +
  aes(shape = Month) +
  geom_point(aes(colour = Site.name), alpha = 0.4, size = 4) +
  geom_polygon(aes(fill=Site.name), alpha=0.5) +
  geom_point(aes(colour = Site.name), size = 1.5) +
  geom_segment(mapping = arrow_map_CAP_13, size = .5,
               data = arrowdf_CAP_13, color = "gray", arrow = arrowhead_CAP_13) +
  geom_text(mapping = label_map_CAP_13, size = 4,
            data = arrowdf_CAP_13, show.legend = FALSE)+
  scale_color_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(13)) +
  scale_fill_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(13))

ps.b.r.CAP_sig.plot<-plot_grid(CAP_sig_plot_a12, CAP_sig_plot_a13,
                               labels=c("A","B"))
```
```{r, fig.height=8, fig.width=14}
ps.b.r.CAP_sig.plot
```

So, as previously seen, two major groups composed of high-S and high-Fe samples. It seems DO and SO4 drive community structure in opposite ways. Also, temperature seems to play a role in defining high-S sample groups.

Is this ordination's goodness of fit better than random models? i.e. are the constraints in this model significant?

```{r}
anova(CAP_ps_cap_r, perm=1e3)
```

Do these significant metadata variables have the same dispersions?

* Homogeneity of dispersion test for significant categories:

***Null hypothesis**: tested metadata categories have the same dispersions (variance).
```betadisper``` calculates average distances of samples to group centroids (from metadata categories)
```permutest``` calculates ANOVAS on average distances from betadisper() to test if one or more groups is more variable than others

```{r betadisper}
#Site name
beta_s <- betadisper(ps_cap_r_jsd, sampledf_cap$Site.name)
permutest(beta_s)
#Month
beta_m <- betadisper(ps_cap_r_jsd, sampledf_cap$Month)
permutest(beta_m)
#Temperature
beta_t <- betadisper(ps_cap_r_jsd, sampledf_cap$Temp.C)
permutest(beta_t)
#Cl
beta_c <- betadisper(ps_cap_r_jsd, sampledf_cap$Cl..µg.ml)
permutest(beta_c)
#DO
beta_d <- betadisper(ps_cap_r_jsd, sampledf_cap$DO..sat)
permutest(beta_d)
# SO4
beta_so <- betadisper(ps_cap_r_jsd, sampledf_cap$SO4..µg.ml)
permutest(beta_so)
#mn
beta_mn <- betadisper(ps_cap_r_jsd, sampledf_cap$X55.Mn.ng.ml)
permutest(beta_mn)
#Y
beta_y <- betadisper(ps_cap_r_jsd, sampledf_cap$X89.Y.ng.ml)
permutest(beta_y)
```

We can only reject the null hypothesis that site names have the same dispersions (variance). However, this might mean that adonis results can't be trusted for this category.

Ok then, but what microbial groups are influenced by these variables? More ordination on the way.

```{r taxa cap on sig metadata, message=FALSE, warning=FALSE}
ps_cap_r_april = subset_samples(ps_cap_r, Month == "April")
ps_cap_r_april_f = filter_taxa(ps_cap_r_april, function(x) sum(x) > .01, TRUE)
ps_cap_r_august = subset_samples(ps_cap_r, Month == "August")
ps_cap_r_august_f = filter_taxa(ps_cap_r_august, function(x) sum(x) > .01, TRUE)
ps_cap_r_december = subset_samples(ps_cap_r, Month == "December")
ps_cap_r_december_f = filter_taxa(ps_cap_r_december, function(x) sum(x) > .01, TRUE)
#april
CAP_ps_cap_r_april_f <- ordinate(ps_cap_r_april_f,
                           method = "CAP",
                           distance = "jsd",
                           k=3, trymax=1e3, weighted=TRUE,
                        formula = ~ Site.name + Month + Temp.C + DO..sat + Cl..µg.ml + SO4..µg.ml+ X55.Mn.ng.ml + X89.Y.ng.ml + X139.La.ng.ml)

arrowmat_CAP_taxa_april <- vegan::scores(CAP_ps_cap_r_april_f, display = "bp")
arrowmat_CAP_taxa_april<-arrowmat_CAP_taxa_april[!grepl("Site.name", rownames(arrowmat_CAP_taxa_april)),]
arrowmat_CAP_taxa_april<-arrowmat_CAP_taxa_april[!grepl("Month", rownames(arrowmat_CAP_taxa_april)),]
arrowdf_CAP_taxa_april <- data.frame(labels = rownames(arrowmat_CAP_taxa_april), arrowmat_CAP_taxa_april)
arrow_map_CAP_taxa_april <- aes(xend = CAP1, yend = CAP2,
                     x = 0, y = 0,
                     shape = NULL, color = NULL,
                     label = labels)
label_map_CAP_taxa_april <- aes(x = 1.3 * CAP1, y = 1.3 * CAP2,
                     shape = NULL, color = NULL,
                     label = labels)
arrowhead_CAP_taxa_april = arrow(length = unit(0.02, "npc"))

april_phyla=nrow(as.data.frame(table(tax_table(ps_cap_r_april_f)[, "Phylum"], exclude = NULL)))

CAP_april_taxa_sig_plot_a12 <- plot_ordination(ps_cap_r_april_f,
                                    ordination = CAP_ps_cap_r_april_f, 
                                    color = "Phylum",
                                    axes = c(1,2), type="taxa") 
ov_CAP_april_taxa_sig_plot_a12<-subset(CAP_april_taxa_sig_plot_a12$data,
                           Phylum %in% c("Proteobacteria","Omnitrophica","Parcubacteria"))
ov_CAP_april_taxa_sig_plot_a12$Phylum=as.character(ov_CAP_april_taxa_sig_plot_a12$Phylum)

final_CAP_april_taxa_sig_plot_a12 = CAP_april_taxa_sig_plot_a12 +
  geom_point(data = ov_CAP_april_taxa_sig_plot_a12,
             aes(x=CAP1, y=CAP2, shape = Phylum, alpha = 0.6, size = 6)) +
  geom_segment(mapping = arrow_map_CAP_taxa_april, size = .5, 
               data = arrowdf_CAP_taxa_april, color = "black", arrow = arrowhead_CAP_taxa_april) +
  geom_text(mapping = label_map_CAP_taxa_april, size = 5, fontface="bold",
            data = arrowdf_CAP_taxa_april, show.legend = FALSE) +
  scale_color_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(april_phyla),
                     na.value = "black") +
  scale_fill_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(april_phyla),
                    na.value = "black")  +
  guides(fill = guide_legend(nrow = april_phyla/2),
         color = guide_legend(nrow= april_phyla/2),
         size = FALSE,
         shape =FALSE,
         alpha = FALSE)

#august
CAP_ps_cap_r_august_f <- ordinate(ps_cap_r_august_f,
                           method = "CAP",
                           distance = "jsd",
                           k=3, trymax=1e3, weighted=TRUE,
                        formula = ~ Site.name + Month + Temp.C + DO..sat + Cl..µg.ml + SO4..µg.ml+ X55.Mn.ng.ml + X89.Y.ng.ml + X139.La.ng.ml)

arrowmat_CAP_taxa_august <- vegan::scores(CAP_ps_cap_r_august_f, display = "bp")
arrowmat_CAP_taxa_august<-arrowmat_CAP_taxa_august[!grepl("Site.name", rownames(arrowmat_CAP_taxa_august)),]
arrowmat_CAP_taxa_august<-arrowmat_CAP_taxa_august[!grepl("Month", rownames(arrowmat_CAP_taxa_august)),]
arrowdf_CAP_taxa_august <- data.frame(labels = rownames(arrowmat_CAP_taxa_august), arrowmat_CAP_taxa_august)
arrow_map_CAP_taxa_august <- aes(xend = CAP1, yend = CAP2,
                     x = 0, y = 0,
                     shape = NULL, color = NULL,
                     label = labels)
label_map_CAP_taxa_august <- aes(x = 1.3 * CAP1, y = 1.3 * CAP2,
                     shape = NULL, color = NULL,
                     label = labels)
arrowhead_CAP_taxa_august = arrow(length = unit(0.02, "npc"))

august_phyla=nrow(as.data.frame(table(tax_table(ps_cap_r_august_f)[, "Phylum"], exclude = NULL)))

CAP_august_taxa_sig_plot_a12 <- plot_ordination(ps_cap_r_august_f,
                                    ordination = CAP_ps_cap_r_august_f, 
                                    color = "Phylum",
                                    axes = c(1,2), type="taxa")

ov_CAP_august_taxa_sig_plot_a12<-subset(CAP_august_taxa_sig_plot_a12$data,
                           Phylum %in% c("Proteobacteria","Omnitrophica","Parcubacteria"))
ov_CAP_august_taxa_sig_plot_a12$Phylum=as.character(ov_CAP_august_taxa_sig_plot_a12$Phylum)

final_CAP_august_taxa_sig_plot_a12 = CAP_august_taxa_sig_plot_a12 +
  geom_point(data = ov_CAP_august_taxa_sig_plot_a12,
             aes(x=CAP1, y=CAP2, shape = Phylum, alpha = 0.6, size = 6)) +
  geom_segment(mapping = arrow_map_CAP_taxa_august, size = .5, 
               data = arrowdf_CAP_taxa_august, color = "black", arrow = arrowhead_CAP_taxa_august) +
  geom_text(mapping = label_map_CAP_taxa_august, size = 5, fontface="bold",
            data = arrowdf_CAP_taxa_august, show.legend = FALSE) +
  scale_color_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(august_phyla),
                     na.value = "black") +
  scale_fill_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(august_phyla),
                    na.value = "black")  +
  guides(fill = guide_legend(nrow = august_phyla/2),
         color = guide_legend(nrow=august_phyla/2),
         shape = guide_legend(title="Shape"),
         size = FALSE,
         alpha = FALSE)

#december
CAP_ps_cap_r_december_f <- ordinate(ps_cap_r_december_f,
                           method = "CAP",
                           distance = "jsd",
                           k=3, trymax=1e3, weighted=TRUE,
                        formula = ~ Site.name + Month + Temp.C + DO..sat + Cl..µg.ml + SO4..µg.ml+ X55.Mn.ng.ml + X89.Y.ng.ml + X139.La.ng.ml)

arrowmat_CAP_taxa_december <- vegan::scores(CAP_ps_cap_r_december_f, display = "bp")
arrowmat_CAP_taxa_december<-arrowmat_CAP_taxa_december[!grepl("Site.name", rownames(arrowmat_CAP_taxa_december)),]
arrowmat_CAP_taxa_december<-arrowmat_CAP_taxa_december[!grepl("Month", rownames(arrowmat_CAP_taxa_december)),]
arrowdf_CAP_taxa_december <- data.frame(labels = rownames(arrowmat_CAP_taxa_december), arrowmat_CAP_taxa_december)
arrow_map_CAP_taxa_december <- aes(xend = CAP1, yend = CAP2,
                     x = 0, y = 0,
                     shape = NULL, color = NULL,
                     label = labels)
label_map_CAP_taxa_december <- aes(x = 1.3 * CAP1, y = 1.3 * CAP2,
                     shape = NULL, color = NULL,
                     label = labels)
arrowhead_CAP_taxa_december = arrow(length = unit(0.02, "npc"))

december_phyla=nrow(as.data.frame(table(tax_table(ps_cap_r_december_f)[, "Phylum"], exclude = NULL)))

CAP_december_taxa_sig_plot_a12 <- plot_ordination(ps_cap_r_december_f,
                                    ordination = CAP_ps_cap_r_december_f, 
                                    color = "Phylum",
                                    axes = c(1,2), type="taxa")
ov_CAP_december_taxa_sig_plot_a12<-subset(CAP_december_taxa_sig_plot_a12$data,
                           Phylum %in% c("Proteobacteria","Omnitrophica","Parcubacteria"))
ov_CAP_december_taxa_sig_plot_a12$Phylum=as.character(ov_CAP_december_taxa_sig_plot_a12$Phylum)

final_CAP_december_taxa_sig_plot_a12 = CAP_december_taxa_sig_plot_a12 +
  geom_point(data = ov_CAP_december_taxa_sig_plot_a12,
             aes(x=CAP1, y=CAP2, shape = Phylum, alpha = 0.6, size = 6)) +
  geom_segment(mapping = arrow_map_CAP_taxa_december, size = .5, 
               data = arrowdf_CAP_taxa_december, color = "black", arrow = arrowhead_CAP_taxa_december) +
  geom_text(mapping = label_map_CAP_taxa_december, size = 5, fontface="bold",
            data = arrowdf_CAP_taxa_december, show.legend = FALSE) +
  scale_color_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(december_phyla),
                     na.value = "black") +
  scale_fill_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(december_phyla),
                    na.value = "black")  +
  guides(fill = guide_legend(nrow = december_phyla/2),
         color = guide_legend(nrow= december_phyla/2),
         size = FALSE,
         shape =FALSE,
         alpha = FALSE)

ps_cap_r_month<-plot_grid(final_CAP_april_taxa_sig_plot_a12, final_CAP_august_taxa_sig_plot_a12, final_CAP_december_taxa_sig_plot_a12,
                               labels=c("April","August","December"), align = 'v', ncol = 1,
                               hjust = -1, vjust = -1)

```
```{r, fig.height=15, fig.width=14}
ps_cap_r_month
```