---
title: "SGG - How does filtering the dataset impact classification and ordination?"
author: "André Soares"
date: "28 February 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r initial loadings, include=FALSE}
library(dada2); packageVersion("dada2")
library(ggplot2); packageVersion("ggplot2")
library(phyloseq); packageVersion("phyloseq")
library(data.table); packageVersion("data.table")
library(DESeq2); packageVersion("DESeq2")
library(dplyr)
library(plyr)
library(magrittr)
library(RColorBrewer)
library(phangorn)
library(DECIPHER)
library("structSSI")
library(ips)
library(msa)
library(picante)
library(igraph)
library(cowplot)
library(plotly)
library(lemon)
```

Loading the dataset.

```{r load SV data}
SGG_SVt = readRDS("/media/andre/B2F8C9A0F8C962E9/SGG_16S_analysis/SGG_16S_data_trimmed_DADA2_table/seqtab_final.rds")
SGG_Tax = readRDS("/media/andre/B2F8C9A0F8C962E9/SGG_16S_analysis/SGG_16S_data_trimmed_DADA2_table/tax_final.rds")
sgg_metadata<-read.csv("/media/andre/B2F8C9A0F8C962E9/SGG_16S_analysis/SGG_16S_metadata/SGG_metadata.csv", header=TRUE)
order<-rownames(SGG_SVt)
sgg_metadata_ord<-sgg_metadata[match(order, sgg_metadata$Sample_Code),]
rownames(sgg_metadata_ord) <- sgg_metadata_ord$Sample_Code
ps <- phyloseq(otu_table(SGG_SVt, taxa_are_rows=FALSE),
               sample_data(sgg_metadata_ord), 
               tax_table(SGG_Tax))
ps.b = subset_samples(ps, Sample_type != "Control")
ps.b <- prune_taxa(taxa_sums(ps.b) > 0, ps.b)
#order months
sample_data(ps.b)$Month = factor(sample_data(ps.b)$Month, 
                                 levels = c("April","August","December"))
#to relative abundances
ps.b = subset_taxa(ps.b, Kingdom %in% c("Archaea", "Bacteria"))
ps.b.r <-  transform_sample_counts(ps.b, function(x) {x/sum(x)} ) 
```

Filtering the dataset to hold only taxa represented by over certain thresholds.

```{r}
ps.b.r.no2 <- prune_taxa(taxa_sums(ps.b) > 2, ps.b.r)
ps.b.r.no5 <- prune_taxa(taxa_sums(ps.b) > 5, ps.b.r)
ps.b.r.no10 <- prune_taxa(taxa_sums(ps.b) > 10, ps.b.r)
ps.b.r.no50 <- prune_taxa(taxa_sums(ps.b) > 50, ps.b.r)
ps.b.r.no100 <- prune_taxa(taxa_sums(ps.b) > 100, ps.b.r)
ps.b.r.no200 <- prune_taxa(taxa_sums(ps.b) > 200, ps.b.r)
ps.b.r.0.005 <- prune_taxa(taxa_sums(ps.b.r) > 0.005, ps.b.r)
ps.b.r.0.01 <- prune_taxa(taxa_sums(ps.b.r) > 0.01, ps.b.r)
ps.b.r.0.05 <- prune_taxa(taxa_sums(ps.b.r) > 0.05, ps.b.r)

ps.b.r.no2
ps.b.r.no5
ps.b.r.no10
ps.b.r.no50
ps.b.r.no100
ps.b.r.no200
ps.b.r.0.005
ps.b.r.0.01
ps.b.r.0.05
```

How does this filtering change genus-level classification?

Showing ratios of classified SVs at different ranks.

```{r}
#Phylum
ps.b.r.no2.p<-as.data.frame(table(tax_table(ps.b.r.no2)[, "Phylum"], exclude = NULL))
ps.b.r.no2.ord.p <- ps.b.r.no2.p[order(-ps.b.r.no2.p$Freq),]
na.no2.p <- ps.b.r.no2.ord.p[is.na(ps.b.r.no2.ord.p$Var1),]
1-(na.no2.p$Freq)/sum(ps.b.r.no2.ord.p$Freq)
#Genus
ps.b.r.no2.g<-as.data.frame(table(tax_table(ps.b.r.no2)[, "Genus"], exclude = NULL))
ps.b.r.no2.ord.g <- ps.b.r.no2.g[order(-ps.b.r.no2.g$Freq),]
na.no2.g <- ps.b.r.no2.ord.g[is.na(ps.b.r.no2.ord.g$Var1),]
1-(na.no2.g$Freq)/sum(ps.b.r.no2.ord.g$Freq)

#Phylum
ps.b.r.no5.g<-as.data.frame(table(tax_table(ps.b.r.no5)[, "Phylum"], exclude = NULL))
ps.b.r.no5.ord.g <- ps.b.r.no5.g[order(-ps.b.r.no5.g$Freq),] 
na.no5.g <- ps.b.r.no5.ord.g[is.na(ps.b.r.no5.ord.g$Var1),]
1-(na.no5.g$Freq)/sum(ps.b.r.no5.ord.g$Freq)
#Genus
ps.b.r.no5.g<-as.data.frame(table(tax_table(ps.b.r.no5)[, "Genus"], exclude = NULL))
ps.b.r.no5.ord.g <- ps.b.r.no5.g[order(-ps.b.r.no5.g$Freq),] 
na.no5.g <- ps.b.r.no5.ord.g[is.na(ps.b.r.no5.ord.g$Var1),]
1-(na.no5.g$Freq)/sum(ps.b.r.no5.ord.g$Freq)

#Phylum
ps.b.r.no10.p<-as.data.frame(table(tax_table(ps.b.r.no10)[, "Phylum"], exclude = NULL))
ps.b.r.no10.ord.p <- ps.b.r.no10.p[order(-ps.b.r.no10.p$Freq),] 
na.no10.p <- ps.b.r.no10.ord.p[is.na(ps.b.r.no10.ord.p$Var1),]
1-(na.no10.p$Freq)/sum(ps.b.r.no10.ord.p$Freq)
#Genus
ps.b.r.no10.g<-as.data.frame(table(tax_table(ps.b.r.no10)[, "Genus"], exclude = NULL))
ps.b.r.no10.ord.g <- ps.b.r.no10.g[order(-ps.b.r.no10.g$Freq),] 
na.no10.g <- ps.b.r.no10.ord.g[is.na(ps.b.r.no10.ord.g$Var1),]
1-(na.no10.g$Freq)/sum(ps.b.r.no10.ord.g$Freq)

#Phylum
ps.b.r.no50.p<-as.data.frame(table(tax_table(ps.b.r.no50)[, "Phylum"], exclude = NULL))
ps.b.r.no50.ord.p <- ps.b.r.no50.p[order(-ps.b.r.no50.p$Freq),] 
na.no50.p <- ps.b.r.no50.ord.p[is.na(ps.b.r.no50.ord.p$Var1),]
1-(na.no50.p$Freq)/sum(ps.b.r.no50.ord.p$Freq)
#Genus
ps.b.r.no50.g<-as.data.frame(table(tax_table(ps.b.r.no50)[, "Genus"], exclude = NULL))
ps.b.r.no50.ord.g <- ps.b.r.no50.g[order(-ps.b.r.no50.g$Freq),] 
na.no50.g <- ps.b.r.no50.ord.g[is.na(ps.b.r.no50.ord.g$Var1),]
1-(na.no50.g$Freq)/sum(ps.b.r.no50.ord.g$Freq)

#Phylum
ps.b.r.no100.p<-as.data.frame(table(tax_table(ps.b.r.no100)[, "Phylum"], exclude = NULL))
ps.b.r.no100.ord.p <- ps.b.r.no100.p[order(-ps.b.r.no100.p$Freq),] 
na.no100.p <- ps.b.r.no100.ord.p[is.na(ps.b.r.no100.ord.p$Var1),]
1-(na.no100.p$Freq)/sum(ps.b.r.no100.ord.p$Freq)
#Genus
ps.b.r.no100.g<-as.data.frame(table(tax_table(ps.b.r.no100)[, "Genus"], exclude = NULL))
ps.b.r.no100.ord.g <- ps.b.r.no100.g[order(-ps.b.r.no100.g$Freq),] 
na.no100.g <- ps.b.r.no100.ord.g[is.na(ps.b.r.no100.ord.g$Var1),]
1-(na.no100.g$Freq)/sum(ps.b.r.no100.ord.g$Freq)

#Phylum
ps.b.r.no200.p<-as.data.frame(table(tax_table(ps.b.r.no200)[, "Phylum"], exclude = NULL))
ps.b.r.no200.ord.p <- ps.b.r.no200.p[order(-ps.b.r.no200.p$Freq),] 
na.no200.p <- ps.b.r.no200.ord.p[is.na(ps.b.r.no200.ord.p$Var1),]
1-(na.no200.p$Freq)/sum(ps.b.r.no200.ord.p$Freq)

#Genus
ps.b.r.no200.g<-as.data.frame(table(tax_table(ps.b.r.no200)[, "Genus"], exclude = NULL))
ps.b.r.no200.ord.g <- ps.b.r.no200.g[order(-ps.b.r.no200.g$Freq),] 
na.no200.g <- ps.b.r.no200.ord.g[is.na(ps.b.r.no200.ord.g$Var1),]
1-(na.no200.g$Freq)/sum(ps.b.r.no200.ord.g$Freq)

#Phylum
ps.b.r.0.005.p<-as.data.frame(table(tax_table(ps.b.r.0.005)[, "Phylum"], exclude = NULL))
ps.b.r.0.005.ord.p <- ps.b.r.0.005.p[order(-ps.b.r.0.005.p$Freq),] 
na.0.005.p <- ps.b.r.0.005.ord.p[is.na(ps.b.r.0.005.ord.p$Var1),]
1-(na.0.005.p$Freq)/sum(ps.b.r.0.005.ord.p$Freq)

#Genus
ps.b.r.0.005.g<-as.data.frame(table(tax_table(ps.b.r.0.005)[, "Genus"], exclude = NULL))
ps.b.r.0.005.ord.g <- ps.b.r.0.005.g[order(-ps.b.r.0.005.g$Freq),] 
na.0.005.g <- ps.b.r.0.005.ord.g[is.na(ps.b.r.0.005.ord.g$Var1),]
1-(na.0.005.g$Freq)/sum(ps.b.r.0.005.ord.g$Freq)

#Phylum
ps.b.r.0.01.p<-as.data.frame(table(tax_table(ps.b.r.0.01)[, "Phylum"], exclude = NULL))
ps.b.r.0.01.ord.p <- ps.b.r.0.01.p[order(-ps.b.r.0.01.p$Freq),] 
na.0.01.p <- ps.b.r.0.01.ord.p[is.na(ps.b.r.0.01.ord.p$Var1),]
1-(na.0.01.p$Freq)/sum(ps.b.r.0.01.ord.p$Freq)

#Genus
ps.b.r.0.01.g<-as.data.frame(table(tax_table(ps.b.r.0.01)[, "Genus"], exclude = NULL))
ps.b.r.0.01.ord.g <- ps.b.r.0.01.g[order(-ps.b.r.0.01.g$Freq),] 
na.0.01.g <- ps.b.r.0.01.ord.g[is.na(ps.b.r.0.01.ord.g$Var1),]
1-(na.0.01.g$Freq)/sum(ps.b.r.0.01.ord.g$Freq)

#Phylum
ps.b.r.0.05.p<-as.data.frame(table(tax_table(ps.b.r.0.05)[, "Phylum"], exclude = NULL))
ps.b.r.0.05.ord.p <- ps.b.r.0.05.p[order(-ps.b.r.0.05.p$Freq),] 
na.0.05.p <- ps.b.r.0.05.ord.p[is.na(ps.b.r.0.05.ord.p$Var1),]
1-(na.0.05.p$Freq)/sum(ps.b.r.0.05.ord.p$Freq)
#Genus
ps.b.r.0.05.g<-as.data.frame(table(tax_table(ps.b.r.0.05)[, "Genus"], exclude = NULL))
ps.b.r.0.05.ord.g <- ps.b.r.0.05.g[order(-ps.b.r.0.05.g$Freq),] 
na.0.05.g <- ps.b.r.0.05.ord.g[is.na(ps.b.r.0.05.ord.g$Var1),]
1-(na.0.05.g$Freq)/sum(ps.b.r.0.05.ord.g$Freq)
```

So it seems that as filtering gets stricter, the ratio of phylum-level classifications goes up from 92% to 94% as do genus-level classifications from 19.5% to 24%. This isn't too weird, kind of expected.

How does ordination respond to this filtering?

```{r}
ps.b.r.no2.pcoa  <- ordinate(ps.b.r.no2, "PCoA", distance="jsd", 
                         k=2, trymax=1e3, weighted=TRUE)
ps.b.r.no5.pcoa  <- ordinate(ps.b.r.no5, "PCoA", distance="jsd", 
                         k=2, trymax=1e3, weighted=TRUE)
ps.b.r.no10.pcoa  <- ordinate(ps.b.r.no10, "PCoA", distance="jsd", 
                         k=2, trymax=1e3, weighted=TRUE)
ps.b.r.no50.pcoa  <- ordinate(ps.b.r.no50, "PCoA", distance="jsd", 
                         k=2, trymax=1e3, weighted=TRUE)
ps.b.r.no100.pcoa  <- ordinate(ps.b.r.no100, "PCoA", distance="jsd", 
                         k=2, trymax=1e3, weighted=TRUE)
ps.b.r.no200.pcoa  <- ordinate(ps.b.r.no200, "PCoA", distance="jsd", 
                         k=2, trymax=1e3, weighted=TRUE)
ps.b.r.0.005.pcoa  <- ordinate(ps.b.r.0.005, "PCoA", distance="jsd", 
                         k=2, trymax=1e3, weighted=TRUE)
ps.b.r.0.01.pcoa  <- ordinate(ps.b.r.0.01, "PCoA", distance="jsd", 
                         k=2, trymax=1e3, weighted=TRUE)
ps.b.r.0.05.pcoa  <- ordinate(ps.b.r.0.05, "PCoA", distance="jsd", 
                         k=2, trymax=1e3, weighted=TRUE)

ps.b.r.no2.pcoa.evals<-ps.b.r.no2.pcoa$values$Eigenvalues
ps.b.r.no2.pcoa.evals.a12<-sum(ps.b.r.no2.pcoa.evals[1:2])/sum(ps.b.r.no2.pcoa.evals)
ps.b.r.no2.pcoa.evals.a12.r<-round(ps.b.r.no2.pcoa.evals.a12, digits=5)*100
ps.b.r.no2.pcoa.evals.a12.r

ps.b.r.no5.pcoa.evals<-ps.b.r.no5.pcoa$values$Eigenvalues
ps.b.r.no5.pcoa.evals.a12<-sum(ps.b.r.no5.pcoa.evals[1:2])/sum(ps.b.r.no5.pcoa.evals)
ps.b.r.no5.pcoa.evals.a12.r<-round(ps.b.r.no5.pcoa.evals.a12, digits=5)*100
ps.b.r.no5.pcoa.evals.a12.r

ps.b.r.no10.pcoa.evals<-ps.b.r.no10.pcoa$values$Eigenvalues
ps.b.r.no10.pcoa.evals.a12<-sum(ps.b.r.no10.pcoa.evals[1:2])/sum(ps.b.r.no10.pcoa.evals)
ps.b.r.no10.pcoa.evals.a12.r<-round(ps.b.r.no10.pcoa.evals.a12, digits=5)*100
ps.b.r.no10.pcoa.evals.a12.r

ps.b.r.no50.pcoa.evals<-ps.b.r.no50.pcoa$values$Eigenvalues
ps.b.r.no50.pcoa.evals.a12<-sum(ps.b.r.no50.pcoa.evals[1:2])/sum(ps.b.r.no50.pcoa.evals)
ps.b.r.no50.pcoa.evals.a12.r<-round(ps.b.r.no50.pcoa.evals.a12, digits=5)*100
ps.b.r.no50.pcoa.evals.a12.r

ps.b.r.no100.pcoa.evals<-ps.b.r.no100.pcoa$values$Eigenvalues
ps.b.r.no100.pcoa.evals.a12<-sum(ps.b.r.no100.pcoa.evals[1:2])/sum(ps.b.r.no100.pcoa.evals)
ps.b.r.no100.pcoa.evals.a12.r<-round(ps.b.r.no100.pcoa.evals.a12, digits=5)*100
ps.b.r.no100.pcoa.evals.a12.r

ps.b.r.no200.pcoa.evals<-ps.b.r.no200.pcoa$values$Eigenvalues
ps.b.r.no200.pcoa.evals.a12<-sum(ps.b.r.no200.pcoa.evals[1:2])/sum(ps.b.r.no200.pcoa.evals)
ps.b.r.no200.pcoa.evals.a12.r<-round(ps.b.r.no200.pcoa.evals.a12, digits=5)*100
ps.b.r.no200.pcoa.evals.a12.r

ps.b.r.0.005.pcoa.evals<-ps.b.r.0.005.pcoa$values$Eigenvalues
ps.b.r.0.005.pcoa.evals.a12<-sum(ps.b.r.0.005.pcoa.evals[1:2])/sum(ps.b.r.0.005.pcoa.evals)
ps.b.r.0.005.pcoa.evals.a12.r<-round(ps.b.r.0.005.pcoa.evals.a12, digits=5)*100
ps.b.r.0.005.pcoa.evals.a12.r

ps.b.r.0.01.pcoa.evals<-ps.b.r.0.01.pcoa$values$Eigenvalues
ps.b.r.0.01.pcoa.evals.a12<-sum(ps.b.r.0.01.pcoa.evals[1:2])/sum(ps.b.r.0.01.pcoa.evals)
ps.b.r.0.01.pcoa.evals.a12.r<-round(ps.b.r.0.01.pcoa.evals.a12, digits=5)*100
ps.b.r.0.01.pcoa.evals.a12.r

ps.b.r.0.05.pcoa.evals<-ps.b.r.0.05.pcoa$values$Eigenvalues
ps.b.r.0.05.pcoa.evals.a12<-sum(ps.b.r.0.05.pcoa.evals[1:2])/sum(ps.b.r.0.05.pcoa.evals)
ps.b.r.0.05.pcoa.evals.a12.r<-round(ps.b.r.0.05.pcoa.evals.a12, digits=5)*100
ps.b.r.0.05.pcoa.evals.a12.r
```

Well, the percentage of variation explained by ordination rises from ~34% to ~39% as filtering goes by. But nothing much really.

But does this change much in the ordination outlook?

```{r}
ps.b.r.0.05.pcoa.evals.a13<-sum(ps.b.r.0.05.pcoa.evals[c(1,3)])/sum(ps.b.r.0.05.pcoa.evals)
ps.b.r.0.05.pcoa.evals.a13.r<-round(ps.b.r.0.05.pcoa.evals.a13, digits=5)*100

ps.b.r.0.05.pcoa.plot.a12 <- plot_ordination(ps.b.r.0.05, ps.b.r.0.05.pcoa, 
                                              axes = c(1,2),
                                       color="Site.name", shape="Month") +
  geom_polygon(aes(fill=Site.name), alpha=0.5) +
  geom_point(alpha=0.5, size=4)+
  scale_alpha(guide = 'none') +
  theme(legend.position = "none") +
  annotate("text", y= 0.3, x =0.25,label=paste("Axes 1 and 2:\n",
                                               ps.b.r.0.05.pcoa.evals.a12.r,"% variation")
           ,hjust=0.5, size=5) +
  scale_color_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(14)) +
  scale_fill_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(14))

ps.b.r.0.05.pcoa.plot.a13 <- plot_ordination(ps.b.r.0.05, ps.b.r.0.05.pcoa, 
                                              axes = c(1,3),
                                       color="Site.name", shape="Month") +
  geom_polygon(aes(fill=Site.name), alpha=0.5) +
  geom_point(alpha=0.5, size=4) +
  scale_alpha(guide = 'none')+
  annotate("text", y= 0.2, x =0.25,label=paste("Axes 1 and 3:\n",
                                               ps.b.r.0.05.pcoa.evals.a13.r,"% variation")
           ,hjust=0.5, size=5) +
  scale_color_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(14)) +
  scale_fill_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(14))

ps.b.r.0.05.pcoa.plot<-plot_grid(ps.b.r.0.05.pcoa.plot.a12, ps.b.r.0.05.pcoa.plot.a13, 
                            labels=c("A","B"))
```
```{r, fig.width=16}
ps.b.r.0.05.pcoa.plot
```

Looks like they're pretty much the same!

However, are genus-level interactions impacted by this filtering and general improvement on the ratios of classification at this rank level?

```{r}
ps.b.r.0.05.g = subset_taxa(ps.b.r.0.05, 
                            Phylum=="Proteobacteria" | Phylum=="Parcubacteria" | Phylum=="Omnitrophica" | Phylum=="Woesearchaeota_(DHVEG-6)")
ps.b.r.0.05.g.glom <- tax_glom(ps.b.r.0.05.g, taxrank = 'Genus', NArm = FALSE)
ps.b.r.0.05.g.glom.df <- data.table(psmelt(ps.b.r.0.05.g.glom))

ps.b.r.0.05.g.glom.df.g<-na.omit(ps.b.r.0.05.g.glom.df, cols=c("Genus"))
ps.b.r.0.05.g.glom.df.g<-subset(ps.b.r.0.05.g.glom.df.g, 
                                                  Genus %in% c("Sideroxydans","Gallionella","Ferriphaselus",
                                                               "Sulfurimonas","Sulfurovum","Sulfuricurvum"))
ps.b.r.0.05.g.glom.df.g$Genus = factor(ps.b.r.0.05.g.glom.df.g$Genus, 
                                                       levels = c("Gallionella","Sideroxydans","Ferriphaselus",
                                                               "Sulfuricurvum","Sulfurovum","Sulfurimonas"))
ps.b.r.0.05.g.glom.df.g$Month = factor(ps.b.r.0.05.g.glom.df.g$Month,
                                           levels=c("December", "August", "April"))
ps.b.r.0.05.g.glom.df.g.plot <- ggplot(ps.b.r.0.05.g.glom.df.g, 
       aes(x=Month, y=Abundance, color=Site.name)) +
  geom_point(aes(fill=Site.name, 
                 color=Site.name)) +
  geom_smooth(aes(group=Site.name),
              method="loess", se = FALSE) +
  facet_wrap(~Genus)+
  theme(axis.title.y = element_blank()) +
  coord_flip() +
  scale_x_discrete(expand=c(0.05, 0.05)) +
  scale_color_manual(values = colorRampPalette(brewer.pal(12, "Paired"))(14))
```
```{r message=FALSE, warning=FALSE, fig.width=16}
ps.b.r.0.05.g.glom.df.g.plot
```

Meh, same crap with slightly better classification ratios.

Not really looking much further into it.